Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-08-08 decorrelation/decorrelation.tex
cleaned up figures and file structure
A few paper updates

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=}
\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

@@

%-------------Header begins here----------------------------------------
\begin{document}
\tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis}

@@

\author{Kevin C. Young}
\email[Corresponding author: ]{kyoung@sandia.gov}
\affiliation{Sandia National Laboratories, Livermore, CA}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006}. These results show that quantum computation is feasible, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

There are many ways in which device noise can fail to be Markovian,\cite{Kelly2018, BlumeKohout2017}, and even when the noise \textit{is} Markovian, many models make simplifying assumptions about the structure of the noise. For instance, Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit compilation to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown and potentially channel dependent error, coherent errors can be averaged away into incoherent error. Indeed, some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise. However, this benefit comes at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing.

@@

\begin{figure}
  \centering
  \includegraphics[width=.75\columnwidth]{simple_example.png}
  \caption{THIS FIGURE SHOULD BE REPLACED}
  \label{fig:simple_example}
\end{figure}
\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BCSs can be done in a variety of ways, using any quantum optimal control technique \cite{Caneva2011, Machnes2018} to find a family of controls. For simplicity in this paper we use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $u_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}u_{ij}H_{j}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U_n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U_n\}$.

@@

\begin{figure*}
\centering
\begin{subfigure}[t]{.5\linewidth}
\includegraphics[width=\textwidth]{1q0.png}
\caption{This is one example of a 1D slice varying over one control.}
\end{subfigure}%
~
\begin{subfigure}[t]{.5\linewidth}
\includegraphics[width=\textwidth]{1q1.png}
\caption{This is another example of a 1D slice varying over one control.}
\end{subfigure}
  \label{fig:1qnum}
\end{figure*}

\begin{figure*}
\centering
\begin{subfigure}[t]{.5\linewidth}
\includegraphics[width=\textwidth]{control_dpn_all0.png}
\caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
\end{subfigure}%
~
\begin{subfigure}[t]{.5\linewidth}
\includegraphics[width=\textwidth]{control_dpn_all2.png}
\caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
\end{subfigure}
  \label{fig:2qnum}
\end{figure*}
\section{Numerical Results}\label{numerical}
\subsection{1Q Gates}\label{1Q Gates}
 In this section, we present numerical results on generating one-qubit gates that together with $RZ(\theta$) rotations are universal for one-qubit computation. Our control Hamiltonian is given as: 
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon \sim \delta \sim \mathcal{N}(0, .001)$ We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}. In our simulation we chose an total evolution time of $T=\pi$, and a number of steps $N=100$, with a threshold infidelity of $1E-3$.

@@

\section{Acknowledgements}

\bibliography{decorrelation.bib}
\end{document}

>>>>

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
% \usepackage[justification=justified,font=small]{caption}
\graphicspath{{./figures/}}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=Decorrelating Errors}
\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

@@

%-------------Header begins here----------------------------------------
\begin{document}
% \tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis}

@@

\author{Kevin C. Young}
% \email[Corresponding author: ]{kyoung@sandia.gov}
\affiliation{Sandia National Laboratories, Livermore, CA}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006} \kcy{more recent citations}. These results show that quantum computation is feasible in principle, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

There are many ways in which device noise can fail to be Markovian,\cite{Kelly2018, BlumeKohout2017}, and even when the noise \textit{is} Markovian, many models make simplifying assumptions about the structure of the noise. For instance, Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but produces overly-pessimistic bounds on device performance.

Other authors have addressed these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit compilation to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown and potentially channel dependent error, coherent errors can be averaged away into incoherent error. Indeed, some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise. However, this benefit comes at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{simple_example.pdf}
  \caption{An example of a balanced control solution. Using optimal control, two implementations of a $Z_\pi$ gate are designed to have equal and opposite sensitivity to errors (if one implementation over-rotates by angle $\theta$, then the other \emph{under}-rotates by $\theta$). Each time the gate is used, one of these implementations is chosen at random. The resulting quantum channel is equivalent to a perfect implementation of the gate followed by dephasing of $\order{\theta^2}$.}
  \label{fig:simple_example}
\end{figure}

@@

\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BCSs can be done in a variety of ways, using any quantum optimal control technique \cite{Caneva2011, Machnes2018} to find a family of controls. For simplicity in this paper we use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $u_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}u_{ij}H_{j}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U_n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U_n\}$.

@@

% \begin{figure*}
% \centering
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{1q0.png}
% \caption{This is one example of a 1D slice varying over one control.}
% \end{subfigure}%
% ~
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{1q1.png}
% \caption{This is another example of a 1D slice varying over one control.}
% \end{subfigure}
%   \label{fig:1qnum}
% \end{figure*}

% \begin{figure*}
% \centering
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all0.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}%
% ~
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all2.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}
%   \label{fig:2qnum}
% \end{figure*}
\section{Numerical Results}\label{numerical}
\subsection{1Q Gates}\label{1Q Gates}
 In this section, we present numerical results on generating one-qubit gates that together with $RZ(\theta$) rotations are universal for one-qubit computation. Our control Hamiltonian is given as:
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon \sim \delta \sim \mathcal{N}(0, .001)$ We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}. In our simulation we chose an total evolution time of $T=\pi$, and a number of steps $N=100$, with a threshold infidelity of $1E-3$.

@@

\section{Acknowledgements}
Sandia National Laboratories is a multimission laboratory managed and operated by National Technology and Engineering Solutions of Sandia, LLC, a wholly owned subsidiary of Honeywell International, Inc., for the U.S. Department of Energy's National Nuclear Security Administration under contract DE-NA0003525.
\bibliography{decorrelation.bib}
\end{document}
